---
title: "Project 1 Report"
author: "Tim Cook, Alex Poplawski, and T.J. Seislove"
date: "3/25/2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(janitor)
```
## Fuel Performance Analysis
### Report By: Tim Cook, T.J. Seislove, and Alex Poplawski


### Introduction

The dataset for this project contains measurements on fuel efficiency for various vehicles, with the goal of comparing real-world efficiency to estimated efficiency. The data was collected by The city of New York and is accessible at data.gov. 
There are 106 observations in the dataset, with each representing all vehicles of a specific year, make, and model. The "vehicle_count" variable indicates how many individual vehicles make up the observation. For example, the 2008 Ford Focus row contains the total information for 49 cars.

Each row records the total mileage driven and gallons of fuel consumed by all vehicles of that type. These are used to calculate the real-world fuel economy. The table also contains the EPA's estimated fuel efficiency for that vehicle type and the expected amount of fuel used given the total mileage and the estimated fuel efficiency All of these variables allow comparisons between actual and expected fuel efficiency.

In addition to the numerical variables, the dataset also has some categorical variables that give additional information about the type of vehicle, namely whether it is a hybrid or not (hybrid_non_hybrid) and the categorization (standard_type) such as van, sedan, or pickup.

Finally, the dataset contains estimated total fuel cost (estimated_fuel_costs_per_gallon) and estimated cost per mile (fuel_costs_per_mile). These are numeric values, but were stored as strings with the $ character proceeding the numbers.

*Dataset Link*

https://catalog.data.gov/dataset/real-world-fuel-efficiency

*Dataset Information*

https://data.cityofnewyork.us/Environment/Real-World-Fuel-Efficiency/mn2p-34if/about_data

### Data Import

```{r}
library(readr)
library(dplyr)

Real_World_Fuel_Efficiency <- read_csv("Real-World_Fuel_Efficiency.csv") %>%
  janitor::clean_names()
```

Our dataset was stored into a dataframe titled, "Real_World_Fuel_Efficiency," which was imported from the CSV file, "Real-World_Fuel_Efficiency.csv" using the read_csv() command from the readr package. We then applied the clean_names() command from the janitor package to convert the variable names to a more consistent, friendlier format with underscores to represent spaces. Doing this allows for an easier analysis process and reduces the likelihood of potential syntax errors in our future work with our analysis.

