---
title: "Project 1 Report"
author: "Tim Cook, Alex Poplawski, and T.J. Seislove"
date: "3/25/2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(janitor)
```
## Fuel Performance Analysis
### Report By: Tim Cook, T.J. Seislove, and Alex Poplawski


### Introduction

The dataset for this project contains measurements on fuel efficiency for various vehicles, with the goal of comparing real-world efficiency to estimated efficiency. The data was collected by The city of New York and is accessible at data.gov. 
There are 106 observations in the dataset, with each representing all vehicles of a specific year, make, and model. The "vehicle_count" variable indicates how many individual vehicles make up the observation. For example, the 2008 Ford Focus row contains the total information for 49 cars.

Each row records the total mileage driven and gallons of fuel consumed by all vehicles of that type. These are used to calculate the real-world fuel economy. The table also contains the EPA's estimated fuel efficiency for that vehicle type and the expected amount of fuel used given the total mileage and the estimated fuel efficiency All of these variables allow comparisons between actual and expected fuel efficiency.

In addition to the numerical variables, the dataset also has some categorical variables that give additional information about the type of vehicle, namely whether it is a hybrid or not (hybrid_non_hybrid) and the categorization (standard_type) such as van, sedan, or pickup.

Finally, the dataset contains estimated total fuel cost (estimated_fuel_costs_per_gallon) and estimated cost per mile (fuel_costs_per_mile). These are numeric values, but were stored as strings with the $ character proceeding the numbers.

*Dataset Link*

https://catalog.data.gov/dataset/real-world-fuel-efficiency

*Dataset Information*

https://data.cityofnewyork.us/Environment/Real-World-Fuel-Efficiency/mn2p-34if/about_data

### Data Import

```{r}
library(readr)
library(dplyr)

Real_World_Fuel_Efficiency <- read_csv("Real-World_Fuel_Efficiency.csv") %>%
  janitor::clean_names()
```

Our dataset was stored into a dataframe titled, "Real_World_Fuel_Efficiency," which was imported from the CSV file, "Real-World_Fuel_Efficiency.csv" using the read_csv() command from the readr package. We then applied the clean_names() command from the janitor package to convert the variable names to a more consistent, friendlier format with underscores to represent spaces. Doing this allows for an easier analysis process and reduces the likelihood of potential syntax errors in our future work with our analysis.

### Conclusion

For our univariate data section, the average city fuel efficiency rating is approximately 23.25 miles per gallon, indicating a moderate level of fuel efficiency. Ratings range from 9 mpg to 55 mpg, highlighting variability among different vehicles. The median rating of 19 mpg suggests a positively skewed distribution, as shown in the histogram, which peaks around 15 mpg with a long right tail, indicating that many vehicles perform below the average. When analyzing the distribution between hybrid and non-hybrid vehicles, we find that out of 106 total vehicles, 34 are hybrids (about 32%), while 72 are non-hybrids (68%). The bar chart clearly shows that non-hybrid vehicles significantly outnumber the hybrid vehicles in our dataset. Finally, the actual fuel economy also displays a wide range, from 5.35 mpg to 54.58 mpg. However, the box and whisker plot indicates a potential outlier, signaling that there are vehicles in our dataset that achieve a higher fuel efficiency. This outlier could skew our data which could indicate why our dataset is positively skewed.

When then looked at the fuel economy by engine type. The average mpg for hybrid vehicles is about 33.68 mpg. But, there is a significant difference compared to the mean mpg for non-hybrid vehicles, which is approximately 13.37 mpg. Looking at our boxplot, hybrid vehicles have a wider range and IQR with the median indicating that many hybrids in the dataset perform well. For non-hybrid vehicles, the box is more narrow with shorter whiskers as well so there is less variability for them as well. Looking at the box and whisker plot and comparing the means, the analysis shows that hybrid vehicles perform better than non-hybrid vehicles in terms of fuel economy. 

We then looked at the total fuel consumed and miles driven between the two vehicle types. The curves indicate that non-hybrid cars consume more fuel per mile driven than hybrid cars. This is consistent with our earlier findings that hybrid cars have better average fuel economy. The correlation between fuel consumed and miles driven is 0.94 for hybrid cars, which indicates that the two variables are very strongly related. For non-hybrids, the correlation is only 0.74. This indicates a strong relationship, but not nearly as strong as the hybrid vehicles. 

This lower correlation seems to be caused in part by the outlier that is the 2015 non-hybrid Toyota Prius. This vehicle totaled over 2.1 million miles, which is more than three times the total mileage of the next highest non-hybrid, the 2015 Ford Escape. However, the Prius in question only consumed 44,874 gallons of fuel. That's only about 1.1 times as much fuel as the Escape. With this in mind, the 2015 Prius non-hybrid appears to have a significant impact on the model, accounting for at least some of the decrease in correlation for non-hybrid cars as compared to hybrids.
